---
title: "trading with different volatilities"
author: 
  - "Gurkamal Pannu"
  - "Maria Nikitanova"
  - "Max Yamamoto"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


```{r functions}
library(tidyverse)
library(moments)

get_d1 <- function(s, k, r, sigma, t) {
  d1 <-
    log(s * exp(r*t) / k) / (sigma * sqrt(t)) + 
       (sigma * sqrt(t)) / 2
  return(d1)
}

bs_call <-  function(s, k, r, sigma, t) {
  d1 <- get_d1(s, k, r,  sigma, t)
  d2 <- d1 - sigma * sqrt(t)
  c0 <-  s * pnorm(d1) - exp(-r * t) * k * pnorm(d2)
  return(c0)
}

get_gamma <- function(s, k_var, r_var, sigma_var, extime){
  d_1 <- get_d1(s, k_var, r_var, sigma_var, extime)
  return(exp(-(d_1^2)/2)/(s * sigma_var * sqrt(extime) * sqrt(2*pi)))
}

get_random_and_stock <-
  function(seed,
           steps,
           period,
           mu_annual,
           sigma_annual,
           r_annual,
           attempt,
           s,
           vol_hedge_annual,
           vol_implied_annual) {
  
  time <- steps / period
  mu <- exp(log(1 + mu_annual)/period) - 1
  sigma <- sigma_annual / sqrt(period)
  r <- exp(log(1 + r_annual) / period) - 1
  vol_hedge <- vol_hedge_annual / sqrt(period)
  vol_implied <- vol_implied_annual / sqrt(period)
  
  set.seed(seed)
  
  rand_variables_raw <- 
    rnorm(attempt * steps, mean = 0, sd = 1) |> 
    matrix(ncol = attempt) |> 
    data.frame() |> 
    tibble() 
  
  rand_variables <-
    rep(s, attempt) |>
    rbind(rand_variables_raw |>
            mutate_all( ~ cumprod(exp((mu-(sigma^2)/2) + sigma * .))) |>
            mutate_all( ~ . * s)) |>
    mutate(extime = time - (row_number() - 1) / period,
           periods = row_number() - 1) |>
    relocate(periods, extime)
  
  stock_path_long <- 
    rand_variables |> 
    select(periods, starts_with("X")) |> 
    pivot_longer(cols = -periods) |> 
    mutate(type = "stock")
  
  trading_simu <- 
    rand_variables |> 
    mutate(
      k_var = k,
      r_var = r_annual,
      sigma_var = sigma_annual,
      t_var = time,
      steps_var = steps,
      vol_hedge_var = vol_hedge_annual,
      vol_implied_var = vol_implied_annual
    ) 
  
  option_path_long <- 
    trading_simu |> 
    mutate_at(vars(starts_with("X")),
              ~ bs_call(., k_var, r_var, vol_implied_var, extime)) |> 
    select(periods, starts_with("X")) |> 
    pivot_longer(cols = -periods) |> 
    mutate(type = "option")
    
  hedge_path_long <- 
    trading_simu |> 
    mutate_at(vars(starts_with("X")),
              ~ pnorm(get_d1(., k_var, r_var, vol_hedge_var, extime))) |>
    
    select(periods, starts_with("X")) |> 
    pivot_longer(cols = -periods) |> 
    mutate(type = "hedge")
    
  gamma_path_long <- 
    trading_simu |> 
    mutate_at(vars(starts_with("X")),
              ~ get_gamma(., k_var, r_var, vol_hedge_var, extime)) |> 
    select(periods, starts_with("X")) |> 
    pivot_longer(cols = -periods) |> 
    mutate(type = "gamma")
  
    dat_all_long <- 
      rbind(stock_path_long,
            option_path_long, 
            hedge_path_long,
            gamma_path_long)
    
    return(dat_all_long)
}

trading_simulation <- function(dat_all_long, lag_no = 1){
  short_option_pl <- 
    dat_all_long |> 
    pivot_wider(values_from = value, names_from = type) |> 
    group_by(name) |> 
    mutate(option_pl = option - lag(option, lag_no),
           hedge_pl = (stock - lag(stock, lag_no)) * lag(hedge, lag_no),
           total_pl = hedge_pl - option_pl) |> 
    drop_na()
  
  trading_results <- 
    short_option_pl |>
    mutate(option_pl = cumsum(option_pl),
           hedge_pl = cumsum(hedge_pl),
           total_pl = cumsum(total_pl)) |> 
    filter(periods == max(periods)) |> 
    select(number = name, total_pl, option_pl, total_pl)
  
  return(trading_results)
}

results_hist <- function(trading_results, x_axis, title){
  trading_results |> 
    select(number, all_of(x_axis)) |> 
    pivot_longer(cols = -number) |> 
    ggplot() +
    aes(x = value) +
    geom_histogram() +
    labs(y = "", x = x_axis, title = title)
}

visualize <- function(dat_long, title){
  dat_long |> 
    ggplot() +
    aes(x = periods, y = value, color = name) + 
    geom_line() +
    theme(legend.position = "none") + 
    labs(y = "", x = "steps", title = title )
}
```

# Next steps (daily simulation)

- Base case actual = implied = hedging = .25 sigma
- Case 1: actual = hedging = .25 but implied = .35.
- Case 2: actual = hedging  = .25 but implied = .15 
- Case 3: hedging = implied = .25 but actual = .35 
- Case 4: hedging = implied = .25 but actual = .15 

the stochastic process of the simulation
$$
S_{t+\Delta t} = S_t \times exp((\mu - \frac{\sigma^2}{2})\Delta t + \sigma \sqrt{\Delta t} N(0,1))
$$

# Case 1
```{r}
attempt <- 1000
period <- 360
steps <- 90
mu_annual <- 0.01 
sigma_annual <- 0.25
r_annual <- 0.01 
s <- 50
k <- 50
seed <- 10

vol_hedge_annual = 0.25
vol_implied_annual = 0.35

dat_all_long <-
  get_random_and_stock(seed,
                       steps,
                       period,
                       mu_annual,
                       sigma_annual,
                       r_annual,
                       attempt,
                       s,
                       vol_hedge_annual,
                       vol_implied_annual)

trading_results <- trading_simulation(dat_all_long, lag_no = 1)

results_hist(trading_results, "total_pl", "Case 1: higher implied volatility")

```

# Case 2

```{r}
vol_implied_annual <- 0.15

dat_all_long_2 <-
  get_random_and_stock(seed,
                       steps,
                       period,
                       mu_annual,
                       sigma_annual,
                       r_annual,
                       attempt,
                       s,
                       vol_hedge_annual,
                       vol_implied_annual)

trading_results_2 <- trading_simulation(dat_all_long_2, lag_no = 1)

results_hist(trading_results_2, "total_pl", "Case 2: lower implied volatility")
```

# summary
```{r}
results <- 
  rbind(trading_results |> mutate(implied = "higher implied vol"),
        trading_results_2 |> mutate(implied = "lower implied vol"))

results |> 
  group_by(implied) |> 
  summarise(sample_average = mean(total_pl),
            sample_sd = sd(total_pl),
            sample_skew = skewness(total_pl),
            sample_kurt = kurtosis(total_pl),
            sample_sharpe = mean(total_pl) / sd(total_pl)) |> 
  #t() |> 
  knitr::kable(digits = 2)
```

## check the stats of simulations
```{r}
dat_all_long_2 |> 
  filter(periods == max(periods), type != "gamma") |> 
  group_by(type) |> 
  summarise(sample_average = mean(value, na.rm = TRUE),
            sample_std = sd(value, na.rm = TRUE)) |> 
  mutate(sample_average = if_else(type == "stock", sample_average/s, sample_average),
         sample_std = if_else(type == "stock", sample_std/s, sample_std)) |> 
  knitr::kable(digits = 2)
```

# Case 3: hedging = implied but looking at actual vol being both lower
```{r}
sigma_annual <- 0.35
vol_hedge_annual = 0.25
vol_implied_annual = 0.25

dat_all_long_3 <-
  get_random_and_stock(seed,
                       steps,
                       period,
                       mu_annual,
                       sigma_annual,
                       r_annual,
                       attempt,
                       s,
                       vol_hedge_annual,
                       vol_implied_annual)

trading_results_3 <- trading_simulation(dat_all_long_3, lag_no = 1)

results_hist(trading_results_3, "total_pl", "Case 3: higher actual volatility")

```

# Case 4: hedging = implied but looking at actual vol being lower
```{r}
sigma_annual <- 0.15
vol_hedge_annual = 0.25
vol_implied_annual = 0.25

dat_all_long_4 <-
  get_random_and_stock(seed,
                       steps,
                       period,
                       mu_annual,
                       sigma_annual,
                       r_annual,
                       attempt,
                       s,
                       vol_hedge_annual,
                       vol_implied_annual)

trading_results_4 <- trading_simulation(dat_all_long_4, lag_no = 1)

results_hist(trading_results_4, "total_pl", "Case 4: lower actual volatility")

```

# summary: Case 3 and Case 4
```{r}
results <- 
  rbind(trading_results_3 |> mutate(actual = "higher actual vol"),
        trading_results_4 |> mutate(actual = "lower actual vol"))

results |> 
  group_by(actual) |> 
  summarise(sample_average = mean(total_pl),
            sample_sd = sd(total_pl),
            sample_skew = skewness(total_pl),
            sample_kurt = kurtosis(total_pl),
            sample_sharpe = mean(total_pl) / sd(total_pl)) |> 
  #t() |> 
  knitr::kable(digits = 2)
```


# Base case: all equal
```{r}
sigma_annual <- 0.25
vol_hedge_annual = 0.25
vol_implied_annual = 0.25

dat_all_long_5 <-
  get_random_and_stock(seed,
                       steps,
                       period,
                       mu_annual,
                       sigma_annual,
                       r_annual,
                       attempt,
                       s,
                       vol_hedge_annual,
                       vol_implied_annual)

trading_results_5 <- trading_simulation(dat_all_long_5, lag_no = 1)

results_hist(trading_results_5, "total_pl", "Base Case: all equal")

```
